---
title: "Predictive text modeling"
author: "Luis Terán"
date: "17/08/2020"
output: 
  md_document:
      variant: markdown_github
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Predictive Text Modeling

## Overview

The main objective of this project is to build a predictive text model.The predictive texting consists of a data processed tool that makes it quicker and easier to write text by suggesting words as you type, predictive text can significantly speed up the input process.

In this file a large corpus of text documents is analized to discover the structure in the data and how words are put together in order to create a model of predctions using N-grams. It is shown how is loaded, cleaned, sampled and analized the data provided by Swiftkey from:

- <https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip>


# Data import and sampling

```{r}
suppressPackageStartupMessages(library(dplyr))
library(tidytext)
library(stringi)
suppressPackageStartupMessages(library(tm))
library(RWeka)
suppressPackageStartupMessages(library(wordcloud))
suppressPackageStartupMessages(library(ggplot2))
```

The data is loaded according to the saved path of the files, in this case in "final" folder. The News dataset is needed to be loaded in binary mode ("rb").

```{r}
blogs <- readLines("final/en_US/en_US.blogs.txt", warn = F)
twitter <- readLines("final/en_US/en_US.twitter.txt", warn = F)
```

```{r}
con <- file("final/en_US/en_US.news.txt", open="rb")
news <- readLines(con, encoding="UTF-8")
close(con)
rm(con)
```

First, we estimate size of loaded variables. The results below show that every dataset is over 250 Mb or even higher.

```{r}
blogsSize<-object.size(blogs)
twitterSize<-object.size(twitter)
newsSize<-object.size(news)
```

```{r}
print(blogsSize, units = "Mb")  
```

```{r}
print(twitterSize, units = "Mb")  
```

```{r}
print(newsSize, units = "Mb")  
```

A word count is performed for every row of the files and then added in order to create a histogram of word count in millions of words contained in each file loaded.

```{r}
blogsRowCount<-stri_count_words(blogs)
twitterRowCount<-stri_count_words(twitter)
newsRowCount<-stri_count_words(news)

blogsCount <- sum(blogsRowCount)
twitterCount <- sum(twitterRowCount)
newsCount <- sum(newsRowCount)

totalCount <- c(Blogs = blogsCount, Twitter = twitterCount, News = newsCount)*(1/1e6)
barplot(height = totalCount, xlab = "File", ylab = "Milions of words", main = "Number of words in each file", col=rgb(0.2,0.4,0.6,0.6))
```
So we have over 30 million words in each file. The exact amount of words contained in each file is shown below:

```{r}
totalCount
```
A summary is displayed for the variation of word counts in each row for every file loaded. As we see, most of the rows contain a few words, less than 50 words, but there are lines with thousands of words in a single row. Also, there are a few other that do not contain any word.

```{r}
summary(blogsRowCount)
```

```{r}
summary(twitterRowCount)
```

```{r}
summary(newsRowCount)
```

## Data Sampling

We first stablish a seed for reproducible purposes and a percentage of reference for the sampling of the data.

```{r}
set.seed(200)
```

```{r}
percentage<-0.01
```

The sampling made is as shown in the next cell where a vector from 1 to 100 is sampled randomly, getting only the 10% of the data.

```{r}
sample(c(1:100), size=100*0.1, replace =FALSE)
```

For speed reasons in the building process, only 1% of the total dataset was sampled. The same process as the cell before is applied for the blogs, news and twitter dataset:

```{r}
blogs <- blogs[sample(c(1:length(blogs)), size=length(blogs)*percentage,
                      replace=FALSE)]

news <- news[sample(c(1:length(news)), size=length(news)*percentage, 
                    replace =FALSE)]

twitter <- twitter[sample(c(1:length(twitter)), size=length(twitter)*percentage,
                          replace =FALSE)]
```

Afeter that, the sampled data is saved a specific folder for later analysis

```{r}
write.csv(blogs, file = "Sample/blogSample.csv", row.names = FALSE, 
          col.names = FALSE)
write.csv(news, file = "Sample/newsSample.csv", row.names = FALSE, 
          col.names = FALSE)
write.csv(twitter, file = "Sample/twitterSample.csv", row.names = FALSE,
          col.names = FALSE)
```

Unnecesay variables are removed for workspace cleaning and memory optimization.

```{r}
rm(blogsCount,blogsRowCount, blogsSize, newsCount, newsRowCount, newsSize,
   twitterCount, twitterRowCount, twitterSize, percentage, totalCount, 
   blogs, news, twitter)
```

## Data cleaning

Once again, the data is loaded but only the sampled files obtained before and combined into a corpus. Transforming data into "Large Simple Corpus" type in order to make tm_map transformations possible

```{r}
corpus <-Corpus(DirSource("Sample/"), readerControl = list(language="en_US"))
```

From the corpus, numbers, punctuation, and leading and/or trailing whitespace is removed. Also, every string element is transformed to lower case.

```{r, results="hide"}
corpus<-tm_map(corpus, removeNumbers)
corpus <- tm_map(corpus, removePunctuation)
corpus <- tm_map(corpus, stripWhitespace)
corpus <- tm_map(corpus, tolower)
```

A special function is created for removing special characters on the corpus.

```{r, results="hide"}
onlyLetters <- function(x)
          gsub("[^A-Za-z///' ]","" , x ,ignore.case = TRUE)
corpus <- tm_map(corpus, onlyLetters)
```

For removing bad words, a dirty, naughty and obscene bad words list was downloaded to remove those from the corpus. The original list can be found in:

- <https://github.com/LDNOOBW/List-of-Dirty-Naughty-Obscene-and-Otherwise-Bad-Words>

For the purpose of this project, only the english version ("en") was used.

```{r, results="hide"}
badwords <- read.csv("en")
badwords <- badwords$X2g1c
corpus <- tm_map(corpus, removeWords, badwords)
```

Stop words are not removed beacuse that could be the case that those words were the expected prediction


# N-Grams analysis

One of the most effective ways to explore the relationship between words is using N-gram models, in other words, examining which words tend to follow others immediately. This can be done by the frequency of times that a word was followed by another (bigram model), the number of times that a word was followed by two other words (trigram model) and so on. For code saving and for readability, a create "n" gram function model was created.

```{r}
calcNgramModel <- function(mycorpus, N){
  token_delim <- " \\t\\r\\n.!?,;\"()"
  token <- NGramTokenizer(mycorpus, Weka_control(min=N,max=N, 
                                                   delimiters = token_delim))
  data <- data.frame(table(token))
  sort_data <- data[order(data$Freq,decreasing=TRUE),]
  sort_data
}
```

The function was used for creating the desired model by passing the recently clened dataset (corpus) and "n" which is the ngram model expected (number of analysed consecutive words). Also, a histogram of the 20 most frequent ngrams and a wordcloud is displayed. This process is repeated for the unigram, bigram and trigram model.

```{r}
unigramModel <- calcNgramModel(corpus, 1)
head(unigramModel)
```

```{r}
ggplot(data=unigramModel[1:20,], aes(x=reorder(token,Freq), y=Freq)) +
  geom_bar(stat="identity",fill=rgb(0.2,0.4,0.6,0.6), colour="black") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + coord_flip() +
  ggtitle("Unigram Model")
```

```{r}
wordcloud(unigramModel$token,unigramModel$Freq,random.order=FALSE,scale = 
            c(2,0.35),min.freq = 500,
          colors = brewer.pal(8,"Dark2"),max.words=150)
```

```{r}
bigramModel <- calcNgramModel(corpus, 2)
head(bigramModel)
```

```{r}
ggplot(data=bigramModel[1:20,], aes(x=reorder(token,Freq), y=Freq)) +
  geom_bar(stat="identity",fill=rgb(0.2,0.4,0.6,0.6), colour="black") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + coord_flip() +
  ggtitle("Bigram Model")
```

```{r}
wordcloud(bigramModel$token,bigramModel$Freq,random.order=FALSE,scale = 
            c(2,0.35),min.freq = 500,
          colors = brewer.pal(8,"Dark2"),max.words=150)
```

```{r}
trigramModel <- calcNgramModel(corpus,3)
head(trigramModel)
```


```{r}
ggplot(data=trigramModel[1:20,], aes(x=reorder(token,Freq), y=Freq)) +
  geom_bar(stat="identity",fill=rgb(0.2,0.4,0.6,0.6), colour="black") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + coord_flip() +
  ggtitle("Trigram Model")
```


```{r}
wordcloud(trigramModel$token,trigramModel$Freq,random.order=FALSE,scale = 
            c(2,0.35),min.freq = 50,
          colors = brewer.pal(8,"Dark2"),max.words=150)
```

The result is a frequency table of the most common consecutive words in english for 1,2,3 consecutive words. For the next word predictions this could be used as if the reference, selecting the "n"-gram model and searching for "n-1" words, we'll be able to predict the next word.
